---
title: "Assignment 1"
date: "1/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install packages

library(readxl)
library(tidyverse)
```

# Problem 1
### Reading in CSV file
```{r}
bcdata <- read.csv(file = 'bcdata_Assignment1.csv')
head(bcdata)

str(bcdata)
```

```{r}
var.names <- c("age", "bmi", "glucose", "insulin", "HOMA-IR", "leptin", "adiponectin", "resistin", "MCP-1", "Classification")
colnames(bcdata) <- var.names
str(bcdata)
```

### Cleaning data

It looks like the data is clean and the variable names are already at the column heads. The only step I've taken is to make the Classification variable a factor, as it's categorical.

```{r}
bcdata$Classification <- as.factor(bcdata$Classification)
levels(bcdata$Classification) <- c("Healthy Controls", "Breast Cancer Patients")
```

### Creating summary tables 

```{r}
summary_df = summary(bcdata)

knitr::kable(summary_df)
```

# Problem 2

I have created new categories with the variable `bmi_cat` based on the numeric ranges of the variable `bmi`. 

```{r}
set.seed(1000)
bcdata$bmi <- as.numeric(bcdata$bmi)
bmi <- bcdata$bmi

bmi_cat <- numeric()
bmi_cat[bmi < 16.5] <- 1
bmi_cat[bmi < 18.5 & bmi > 16.5] <- 2
bmi_cat[bmi >= 18.5 & bmi < 24.9] <- 3
bmi_cat[bmi >= 25 & bmi < 29.9] <- 4
bmi_cat[bmi >= 30 & bmi < 34.9] <- 5
bmi_cat[bmi >= 35 & bmi < 39.9] <- 6
bmi_cat[bmi >= 40] <- 7

bmi_cat <- as.factor(bmi_cat)
head(bmi_cat)
levels(bmi_cat) <- c("Severely underweight", "Underweight", "Normal weight", "Overweight", "Obesity class I", "Obesity class II", "Obesity class III")
```

I then added this new variable to a column using `mutate`.

```{r}
mutate(bcdata,
       bmi_category = bmi_cat)
```

# Problem 3

```{r}
# library
library(ggplot2)

# dataset
category <- c(rep("Severely underweight", 2) , rep("Underweight" , 2) , rep("Normal weight" , 2) , rep("Overweight" , 2) , rep("Obesity class I", 2) , rep("Obesity class II" , 2) , rep("Obesity class III", 2) )
status <- rep(c("Healthy Controls", "Breast Cancer Patients"), 7)
value <- abs(rnorm(14, 12))
data <- data.frame(category, status, value)

# stacked and proportion
ggplot(data, aes(fill=status, y = value, x = category)) +
  geom_bar(position = "fill", stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5))

```



